---
title: "W09 Task - Interactive Time Series"
author: "Scott Townsend"
date: "`r format(Sys.time(), '%B %d, %Y')`"

execute:
  keep-md: true

format:
  html:
    code-fold: true
    code-line-numbers: true
---

### Libraries

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=12}
library(tidyverse)
library(dygraphs)
library(tidyquant)
```

# Task 

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=12}
stock_symbols <- c("AAPL", "GOOGL", "TSLA")
start_date <- Sys.Date() - years(5)

stock_data <- tq_get(stock_symbols, from = start_date, get = "stock.prices")

stock_returns <- stock_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "daily", type = "log") %>%
  ungroup()

stock_returns_wide <- stock_returns %>%
  select(date, symbol, daily.returns) %>%
  spread(symbol, daily.returns)

dygraph(stock_returns_wide, main = "Daily Stock Returns Over the Last 5 Years") %>%
  dyRangeSelector() %>%
  dyOptions(stackedGraph = FALSE, fillGraph = TRUE) %>%
  dyAxis("y", label = "Daily Return") %>%
  dyLegend(width = 400)
```

# Optional

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=12}
portfolio_returns <- stock_returns %>%
  group_by(date) %>%
  summarize(portfolio_return = mean(daily.returns, na.rm = TRUE)) %>%
  mutate(cumulative_return = cumsum(portfolio_return))

dygraph(portfolio_returns, main = "Cumulative Portfolio Return Over the Last 5 Years") %>%
  dyOptions(stackedGraph = FALSE) %>%
  dyAxis("y", label = "Cumulative Return")
```

# Summary of Findings

Over the past five years, Tesla’s returns were the most unpredictable, with bigger daily changes, while Apple and Google showed steadier, smaller changes. The overall portfolio had a positive trend, especially in the last two years, suggesting that a mix of stable and high-growth tech stocks can provide a good balance of risk and reward